Individual and Differential Harm in Redistricting

### 2025-08-25 paper/harm.qmd
edits

A key feature of our framework is its clear separation of theoretical quantities of interest and empirical strategies for estimating them [@katz2022].
The methods section discusses the challenge of estimating harm and differential harm from observed data.
Recent advances in redistricting simulation, which can generate a large number of random districting plans that respect relevant legal constraints, are critical to this effort [@fifield2020mcmc; @deford2019; @carter2019; @mccartan2020].
These algorithmically-generated sets of districting plans can provide a realistic set of counterfactuals for our measures.

@@

Once we have defined a set of groups, we define the *differential harm* for two groups $G_k$ and $G_l$ as
$$
  \Delta_{G_k-G_l}H(q) \dfeq \E[H(q)\mid I\in G_k] - \E[H(q)\mid I\in G_l].
$$
So $\Delta_{G_k-G_l}H(q)$ will be positive if voters in group $G_k$ are harmed at a higher rate than voters in group $G_l$, indicating that plan $q$ favors $G_l$ over $G_k$.
The estimation of harm by group may require methods like ecological inference [@king1997solution], as discussed in @sec-estimating.

@@

The estimate of differential partisan harm aligns with the expected partisan skew of each plan, with a value of $-3.6\%$ for the Democratic proposal and $2.0\%$ for the Republican proposal.
This means that, for example, an average Democratic voter is $3.6$pp less likely to be misrepresented than their average Republican counter-partisan.
Overall, the Democratic plan has lower average harm, and a lower ranked marginal deviation [@herschlag2020quantifying].

>>>>

A core feature of our framework is its clear separation of theoretical quantities of interest and empirical strategies for estimating them [@katz2022].
The methods section discusses the challenge of estimating harm and differential harm from observed data.
Recent advances in redistricting simulation, which can generate a large number of random districting plans that respect relevant legal constraints, are critical to this effort [@fifield2020mcmc; @deford2019; @carter2019; @mccartan2020].
These algorithmically-generated sets of districting plans can provide a realistic set of counterfactuals for our measures.

@@

Once we have defined a set of groups, we define the *differential harm* for two groups $G_k$ and $G_l$ as
$$
  \Delta_{G_k-G_l}H(q) \dfeq \E[H(q)\mid I\in G_k] - \E[H(q)\mid I\in G_l].
$$
So $\Delta_{G_k-G_l}H(q)$ will be positive if voters in group $G_k$ are harmed at a higher rate than voters in group $G_l$, indicating that plan $q$ favors $G_l$ over $G_k$.
The estimation of harm by group may require methods like ecological inference, as discussed in the methods section.

@@

The estimate of differential partisan harm aligns with the expected partisan skew of each plan, with a value of $-3.6\%$ for the Democratic and $2.0\%$ for the Republican proposal.
This means that, for example, an average Democratic voter is $3.6$pp less likely to be misrepresented than their average Republican counter-partisan.
Overall, the Democratic plan has lower average harm, and a lower ranked marginal deviation [@herschlag2020quantifying].
